Well-quasi-orderings on word languages

### 2024-10-21 src/prefixes.tex
correct knowledge usage

One immediate application of \cref{antichain-branches-prefix:lem} is
that \kl{antichain branches} can be described inside the \kl{tree of prefixes}
by a monadic second order formula (\kl{$\MSO$-formula}), allowing us to
leverage the decidability of $\MSO$ over infinite binary trees \cite[Theorem
1.1]{RAB69}. This result will follow from our general decidability result
(\cref{infix-wqo-is-emptiness:thm}) but is worth stating on its own for its
simplicity.

\begin{corollary}
    \label{prefix-wqo-reg-decidable:cor}
    If $L$ is regular, then the existence of an infinite antichain is decidable.
\end{corollary}
\begin{proof}
    If $L$ is regular, then it is \kl{$\MSO$-definable}, and there 
    exists a formula $\varphi(x)$ in \kl{$\MSO$} that selects nodes 
    of the \kl{tree of prefixes} that belong to $L$. Now, to decide whether there
    exists an \kl{antichain branch} for $L$, we can simply check whether
    the following formula is satisfied:
    \begin{equation*}
        \exists B. 
        B \text{ is a branch } \land
        \forall x \in B, \exists y. y \text{ is a child of } x \land \varphi(y) \land y \not\in B
        \quad .
    \end{equation*}
    Because the above formula is an $\MSO$-formula over the infinite
    $\Sigma$-branching tree, whether it is satisfied is decidable
    as an easy consequence of the decidability of $\MSO$ over infinite binary
    trees
    \cite[Theorem 1.1]{RAB69}.
\end{proof}

@@

For the proof of the following theorem, we will use special notations to
describe the \intro{upwards closure} of a set $S$ for a relation $\preceq$,
which is defined as $\upset[\preceq]{S} \defined \setof{y \in \Sigma^*}{
\exists x \in S. x \preceq y}$. Anticipating the use of the symmetric notion of
\intro{downwards closure}, let us introduce it as follows: $\dwset[\preceq]{S} \defined \setof{y
\in \Sigma^*}{ \exists x \in S. y \preceq x}$. Abusing notations, we will
write $\upset{w}$ and $\dwset{w}$ for the upwards and downwards closure of a
single element $w$, omitting the ordering relation when it is clear from the
context.

>>>>

One immediate application of \cref{antichain-branches-prefix:lem} is
that \kl{antichain branches} can be described inside the \kl{tree of prefixes}
by a monadic second order formula ($\MSO$-formula), allowing us to
leverage the decidability of $\MSO$ over infinite binary trees \cite[Theorem
1.1]{RAB69}. This result will follow from our general decidability result
(\cref{infix-wqo-is-emptiness:thm}) but is worth stating on its own for its
simplicity.

\begin{corollary}
    \label{prefix-wqo-reg-decidable:cor}
    If $L$ is regular, then the existence of an infinite antichain is decidable.
\end{corollary}
\begin{proof}
    If $L$ is regular, then it is $\MSO$-definable, and there 
    exists a formula $\varphi(x)$ in $\MSO$ that selects nodes 
    of the \kl{tree of prefixes} that belong to $L$. Now, to decide whether there
    exists an \kl{antichain branch} for $L$, we can simply check whether
    the following formula is satisfied:
    \begin{equation*}
        \exists B. 
        B \text{ is a branch } \land
        \forall x \in B, \exists y. y \text{ is a child of } x \land \varphi(y) \land y \not\in B
        \quad .
    \end{equation*}
    Because the above formula is an $\MSO$-formula over the infinite
    $\Sigma$-branching tree, whether it is satisfied is decidable
    as an easy consequence of the decidability of $\MSO$ over infinite binary
    trees
    \cite[Theorem 1.1]{RAB69}.
\end{proof}

@@

For the proof of the following theorem, we will use special notations to
describe the \intro{upwards closure} of a set $S$ for a relation $\preceq$,
which is defined as $\upset[\preceq]{S} \defined \setof{y \in \Sigma^*}{
\exists x \in S. x \preceq y}$. Anticipating the use of the symmetric notion of
\intro{downwards closure}, let us introduce it as follows: $\dwset[\preceq]{S} \defined \setof{y
\in \Sigma^*}{ \exists x \in S. y \preceq x}$. Abusing notations, we will
write $\upset{w}$ and $\dwset{w}$ for the upwards and downwards closure of a
single element $w$, omitting the ordering relation when it is clear from the
context. A set $S$ is called \intro{downwards closed} if $\dwset{S} = S$.
